Is essentiality of genes conserved in Enterobacteriaceae?

### 2016-07-15 documents/manuscript.tex
Added SL3261, reorganised everything

\usepackage{graphicx}
\graphicspath{{figs/}}
%------------------------------------------------
% OPTIONAL POSTSCRIPT FONT FILES
%------------------------------------------------

@@

\title{Title of the publication} % For titles, only capitalize the first letter

@@

\author{John Smith\affil{1}{University of California},
James Smith\affil{2}{University of Oregon}
\and
Jane Smith\affil{1}{}}

@@

Although the essentiality of genes has been studied in a variety of organisms, there is still room to study the evolutionary conservation of essentiality. Barquist et al.\@ \cite{barquist_comparison_2013} have used transposon-directed insertion-site sequencing to study the differentiation of the essentiality of genes in \textit{Salmonella} serovars Typhi and Typhimurium which has led to divergence in their pathogenecity and host ranges. We extend this research by studying 12 bacterial strains. These include \textit{Salmonella} enterica subsp.\@ enterica serovar Typhi str.\@ Ty2, \textit{Salmonella} enterica subsp.\@ enterica serovar Enteritidis str.\@ P125109, \textit{Salmonella} enterica subsp.\@ enterica serovar Typhimurium str.\@ SL1344, \textit{Salmonella} enterica subsp.\@ enterica serovar Typhimurium str.\@ A130, \textit{Salmonella} enterica subsp.\@ enterica serovar Typhimurium str.\@ D23580, \textit{Escherichia} coli UPEC ST131, \textit{Escherichia} coli ETEC CS17, \textit{Escherichia} coli ETEC H10407, \textit{Citrobacter} rodentium ICC168, \textit{Klebsiella} pneumoniae RH201207, \textit{Klebsiella} pneumoniae subsp.\@ pneumoniae Ecl8, and \textit{Enterobacter} cloacae subsp.\@ cloacae NCTC 9394. All these strains are selected from Enterobacteriaceae family.

@@

\section{Bias}

\section{Gene classes}
\subsection{Genus specific}
\subsection{Single copy}
\subsection{Multi-copy}
\section{Evolution of essentiality}

\section{Case study of genes}
\subsection{Core genes}
\subsection{Accessory genes}

@@

%----------------------------------------------------------------------------------------
%	FIGURES AND TABLES
%----------------------------------------------------------------------------------------

>>>>

\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\graphicspath{{../figures/}}
%------------------------------------------------
% OPTIONAL POSTSCRIPT FONT FILES
%------------------------------------------------

@@

\title{Is essentiality of genes conserved in Enterobacteriaceae?} % For titles, only capitalize the first letter

@@

\author{Fatemeh Ashari-Ghomi\affil{1}{University of Canterbury},
Paul P. Gardner\affil{1}{}
\and
Lars Barquist\affil{2}{Wurzburg University}}

@@

Although the essentiality of genes has been studied in a variety of organisms, there is still room to study the evolutionary conservation of essentiality. Barquist et al.\@ \cite{barquist_comparison_2013} have used transposon-directed insertion-site sequencing to study the differentiation of the essentiality of genes in \textit{Salmonella} serovars Typhi and Typhimurium which has led to divergence in their pathogenecity and host ranges. We extend this research by studying 13 bacterial strains from Enterobacteriaceae. These strains are depicted in Fig.\@~\ref{fig:species-tree}.

@@

\section{Are transposons biased towards certain positions?}
\begin{itemize}
\item There is a bias towards the position of the gene (Fig.~\ref{fig:distance-bias}).
\item There is a negligible bias towards certain motifs (Fig.~\ref{fig:logos})
\item The G-C bias needs to be studied further (Fig.~\ref{fig:GC-bias})
\item The number of insertions on the 3’ and 5’ ends is more than the internal region in essential genes and less than the internal region in beneficial losses. (Fig.~\ref{fig:insertion-position-bias})
\end{itemize}
\section{Gene classes}
\subsection{Genus specific}
\subsection{Single copy}
\subsection{Multi-copy}
\section{Evolution of essentiality}
\begin{itemize}
\item UpSetR results (Fig.~\ref{fig:upsetr})
\item Stringent
\item Dollo law
\item Ancestral insertion index
\end{itemize}
\section{Case study of genes}
\subsection{Core genes}
\subsection{Accessory genes}

@@

%----------------------------------------------------------------------------------------
%	FIGURES AND TABLES
%----------------------------------------------------------------------------------------
\begin{figure}
\includegraphics[scale=0.3]{phylosift-aa-raxmlbootstrap.pdf}
\caption{The species tree containing the 13 strains under study and Escherichia \textit{coli} K-12 MG1655 studied in Keio collection \cite{baba_construction_2006}. We have generated the tree by running RAxML \cite{stamatakis_raxml_2014} on Phylosift \cite{darling_phylosift:_2014} amino acid markers.}
\label{fig:species-tree}
\end{figure}

\begin{figure*}
\centering
\begin{tabular}{c c c}
\includegraphics[page=2, scale=0.25]{biases.pdf}&
\includegraphics[page=5, scale=0.25]{biases.pdf}&
\includegraphics[page=8, scale=0.25]{biases.pdf}\\
\includegraphics[page=11, scale=0.25]{biases.pdf}&
\includegraphics[page=14, scale=0.25]{biases.pdf}&
\includegraphics[page=17, scale=0.25]{biases.pdf}\\
\includegraphics[page=20, scale=0.25]{biases.pdf}&
\includegraphics[page=23, scale=0.25]{biases.pdf}&
\includegraphics[page=26, scale=0.25]{biases.pdf}\\
\includegraphics[page=29, scale=0.25]{biases.pdf}&
\includegraphics[page=32, scale=0.25]{biases.pdf}&
\includegraphics[page=35, scale=0.25]{biases.pdf}\\
&\includegraphics[page=38, scale=0.25]{biases.pdf}&\\
\end{tabular}
\caption{The plots show the position of the genes within the genome (normalised by the lengths of the genomes) versus the insertion indices of the genes}
\label{fig:distance-bias}
\end{figure*}

\begin{figure*}
\begin{tabular}{c c}
\includegraphics[scale=0.6]{100logo-bits.pdf}&
\includegraphics[scale=0.6]{100logo-prob.pdf}
\end{tabular}
\caption{We have generated the logos from 10 nucleotides flanking the 100 top most frequent insertion sites.}
\label{fig:logos}
\end{figure*}

\begin{figure*}
\centering
\begin{tabular}{c c c}
\includegraphics[page=1, scale=0.25]{biases.pdf}&
\includegraphics[page=4, scale=0.25]{biases.pdf}&
\includegraphics[page=7, scale=0.25]{biases.pdf}\\
\includegraphics[page=10, scale=0.25]{biases.pdf}&
\includegraphics[page=13, scale=0.25]{biases.pdf}&
\includegraphics[page=16, scale=0.25]{biases.pdf}\\
\includegraphics[page=19, scale=0.25]{biases.pdf}&
\includegraphics[page=22, scale=0.25]{biases.pdf}&
\includegraphics[page=25, scale=0.25]{biases.pdf}\\
\includegraphics[page=28, scale=0.25]{biases.pdf}&
\includegraphics[page=31, scale=0.25]{biases.pdf}&
\includegraphics[page=34, scale=0.25]{biases.pdf}\\
&\includegraphics[page=37, scale=0.25]{biases.pdf}&\\
\end{tabular}
\caption{The plots show the G-C contents of the genes (normalised by the lengths of the genes) against their insertion indices}
\label{fig:GC-bias}
\end{figure*}

\begin{figure*}
\begin{tabular}{c c}
\includegraphics[scale=0.5, page=1]{insertion-position-bias.pdf}&
\includegraphics[scale=0.5, page=2]{insertion-position-bias.pdf}\\
\includegraphics[scale=0.5, page=3]{insertion-position-bias.pdf}&
\includegraphics[scale=0.5, page=4]{insertion-position-bias.pdf}
\end{tabular}
\caption{We have divided our genes into 3 segments: 5\% of
the genes on the 5’ end, 20\% of the genes on the 3’ end, and the rest in the middle.}
\label{fig:insertion-position-bias}
\end{figure*}

\begin{figure}
\centering
\includegraphics[scale=0.65, page=1]{upsetr.pdf}\\
\includegraphics[scale=0.65, page=2]{upsetr.pdf}
\caption{The first figure shows the number of core genes between each group of species and the second figure shows the number of core essential genes.}
\label{fig:upsetr}
\end{figure}
